### 2015-06-08 text/ms.Rmd
Edit abstract; add todo

What can we do when multiple ecological models give different impressions of population status? Any single model is unlikely to perform best in all scenarios. However, ensemble methods from the fields climate science and machine learning provide a powerful tool to leverage information across multiple ecological models. Here we evaluate the potential for ensemble methods to improve estimates of fish population status. We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine-learning methods---random forests and boosted regression trees. We build our ensembles on a simulated dataset of 5760 fish stocks and test our models with cross-validation and a global database of `r ram_stocks_n` stock assessments. We find that ensemble models substantially improve estimates of population status and trajectory. Machine-learning ensembles performed the best: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the slope of status. When the simulation-trained ensembles were applied to global fish stocks, the best ensembles still improved all dimensions of performance over the next best individual model. Ensemble models have potential to improve estimates of population status and trajectory across many taxa beyond fish. However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.

@@

We developed and tested ensemble models with both a fully factorial simulated
dataset [@rosenberg2014] and the RAM Legacy Stock Assessment Database
[@ricard2012]. @rosenberg2014 provide a full description of the simulation
model and the code to generate the simulations is available at
<https://github.com/flr/StockSims>. To summarize, the model was built in the
package FLR [@kell2007] for the statistical software \textsf{R} [@r2015]...
included all combinations of the following factors:

>>>>

What can we do when multiple ecological models give different impressions of population status? Any single model is unlikely to perform best in all scenarios. However, ensemble methods from the fields climate science and machine learning provide a powerful tool to leverage information across multiple ecological models. Here we evaluate the potential for ensemble methods to improve estimates of fish population status. We fit four data-limited models of population and exploitation dynamics and combine their estimates of the mean and slope of biomass at maximum sustainable yield (\bbmsy\\) with four ensemble models: a simple average, a linear model, and two machine-learning methods---random forests and boosted regression trees. We build our ensembles on a simulated dataset of 5760 fish stocks and test our models with cross-validation and a global database of `r ram_stocks_n` stock assessments. We find that ensemble models substantially improve estimates of population status and trajectory. Machine-learning ensembles performed the best: accuracy improved `r mean_sim$mach_vs_ind_mare_fold`%, correlation improved from `r mean_sim$ind_corr_range` to `r mean_sim$mach_corr_range`, and bias (median proportional error) declined from `r mean_sim$ind_mre_range` to `r mean_sim$mach_mre_range`. We found similar improvements when predicting the slope of status. When the simulation-trained ensembles were applied to global fish stocks, the best ensembles still improved all dimensions of performance over the next-best individual model. Ensemble models have potential to improve estimates of population status and trajectory across many taxa beyond fish. However, ensemble methods must be carefully tested, formed from a diverse set of accurate models, and built on a dataset sufficiently representative of the populations they are applied to.

@@

We developed and tested ensemble models with both a fully factorial simulated
dataset [@rosenberg2014] and the RAM Legacy Stock Assessment Database
[@ricard2012]. @rosenberg2014 provide a full description of the simulation
model and the code to generate the simulations is available at
<https://github.com/flr/StockSims> (TODO: not there). To summarize, the models were built in the
package FLR [@kell2007] for the statistical software \textsf{R} [@r2015]...
included all combinations of the following factors:
